Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-10-04 paper/paper.Rmd
more changes from @jbracher

\begin{align*}
  \lambda_0 &\sim \text{LogNormal}\left(\log C_0 , 0.025 \times \log C_0 \right)\\
  \lambda_t &= C_{t-1} e^{r_t},\ t > 0  \\
  C_{t} \mid \lambda_{t} &\sim \text{NB}\left(\lambda_t, \phi\right)
\end{align*}

where $r_t$ can be interpreted as the weekly growth rate. $r_t$ is then modelled as a piecewise constant differenced AR(1) process modified such that the dependence of $r_{t-1}$ is multiplied by a decay factor ($\xi_{+,-}$) that varies dynamically according to the sign of $r_{t-1}$.
This assumes that the growth rate is non-stationary with a trend that is independent of the current growth rate (the differenced AR(1) process), the additional decay factor encodes the belief that larger absolute growth rates will tend more quickly towards no growth and that this process may work differently for positive or negative growth rates.
This process can be defined as follows,

@@

In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile [@scoringutils]. 
We also calculate the bias [@Funk2019-vi] of both forecasting approaches, stratified by forecast horizon. This metric aims to capture the tendency for a forecast to under or over-predict. It captures the average proportion of the mass of the forecast distribution that is above or below the true value (and so can range from -1 to 1) with an unbiased forecast having an average bias value of 0.
Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon, to assess the relative difference in performance across the predictive distribution.

@@

## Data summary {-}

@@

```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one and four weeks). 30\\%, 60\\%, and 90\\% prediction intervals are shown. The black line and points are the notified cases as of the date of data extraction rather than those available at the time. b.) A replicate of a.) but with incidence rates on the log scale. c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 12, fig.width = 18}
p <- (
  plot_forecast_custom(forecasts, log = FALSE) |
  plot_forecast_custom(forecasts, log = TRUE) +
  labs(y = "") |
  plot_wis(wis, locs)
) +
  plot_layout(widths = c(2, 2, 1), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre", axis.title = element_text(size = 15)
  )
suppressWarnings(print(p))
```

@@

If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`). When using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`). Across forecast horizons the same pattern of outperformance holds. However, the difference in relative performance was less than when the full probability distribution was accounted for with this becoming more marked as the forecast horizon increased (Figure \@ref(fig:assemble-eval) c).

@@

```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale. a.) The density of the relative score by horizon. Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles. b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon. c.) The average relative weighted interval score and absolute error for the surrogate model compared to the ensemble forecast by forecast horizon. d.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon. The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast. The vertical black lines on the y-axis give individual relative scores."}
p <- (
  (
    (
      plot_rwis_horizon / plot_rwis_by_month
    )  +
    plot_layout(heights = c(4, 5), guides = "collect")
  ) |
  (
    (
      plot_relative_summary(relative_summary) +
      plot_layout(guides = "keep")
    ) +
    plot_rwis_location +
    plot_layout(heights = c(1, 7), guides = "collect")
  )
) +
  plot_layout(guides = "collect", widths = c(1, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre", axis.title = element_text(size = 15)
  )
suppressWarnings(print(p))
```

@@

Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
This behaviour became more prominent as the forecast horizon increased.
In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to over cover for intervals above the median.
At longer horizons, however, quantiles below the median became liable to under cover.

@@

```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon. Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the nominal level of the interval (i.e., 30\\%, 60\\% and 90\\%, respectively). b.) Empirical coverage by quantile for both the surrogate and ensemble forecasts. A well-calibrated forecast should have empirical quantiles that match the theoretical ones. The green area of this figure corresponds to conservative forecasts. c.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast. d.) Bias of the ensemble and surrogate forecasts stratified by horizon."}
p <- (
  (
    (
      plot_coverage_range(coverage, ranges)
    )
  ) | (
    plot_coverage_quantiles(scores) + (
      plot_rel_score_by_quantile(relative_interval_score) +
      plot_bias(summary_performance)
    ) +
    plot_layout(heights = c(2, 1), guides = "collect")
  )
) +
  plot_layout(widths = c(1, 2), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre", axis.title = element_text(size = 15)
  )
suppressWarnings(print(p))
```

>>>>

\begin{align*}
  \lambda_0 &\sim \text{LogNormal}\left(\log C_0 , 0.025 \times \log C_0 \right)\\
  \lambda_t &= C_{t-1} e^{r_t},\ t > 0  \\
  C_{t} \mid \lambda_{t} &\sim \text{NB}\left(\lambda_t, \phi\right)
\end{align*}
where the mean and variance of the negative binomial are defined as \[ \mathbb{E}[n] = \mu \ \ \
\text{ and } \ \ \ \text{Var}[n] = \mu + \frac{\mu^2}{\phi}. \]
Here $r_t$ can be interpreted as the weekly growth rate. $r_t$ is then modelled as a piecewise constant differenced AR(1) process modified such that the dependence of $r_{t-1}$ is multiplied by a decay factor ($\xi_{+,-}$) that varies dynamically according to the sign of $r_{t-1}$.
This assumes that the growth rate is non-stationary with a trend that is independent of the current growth rate (the differenced AR(1) process), the additional decay factor encodes the belief that larger absolute growth rates will tend more quickly towards no growth and that this process may work differently for positive or negative growth rates.
This process can be defined as follows,

@@

In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile [@scoringutils]. 
We also calculate the bias (see [@scoringutils] and [@Funk2019-vi] for a more detailed definition) of both forecasting approaches, stratified by forecast horizon. This metric aims to capture the tendency for a forecast to under or over-predict. It captures the average proportion of the mass of the forecast distribution that is above or below the true value (and so can range from -1 to 1) with an unbiased forecast having an average bias value of 0.
Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon, to assess the relative difference in performance across the predictive distribution.

@@

##  Summary of the European COVID-19 Forecast Hub Setting {-}

@@

```{r vis-forecasts, fig.cap = "a.) Forecasts of notified test positive cases (per 10,000 population) by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom,  by forecast horizon (one and four weeks). 30\\%, 60\\%, and 90\\% prediction intervals are shown. The black line and points are the notified cases as of the date of data extraction rather than those available at the time. b.) A replicate of a.) but with incidence rates on the log scale. c.) Weighted interval scores at the one-week and four-week forecast horizon by epidemiological week in Germany, Greece, Italy, Poland, Slovakia, and the United Kingdom on the log scale.", out.width = "95%", fig.height = 12, fig.width = 18}
p <- (
  plot_forecast_custom(forecasts, log = FALSE) |
  plot_forecast_custom(forecasts, log = TRUE) +
  labs(y = "") |
  plot_wis(wis, locs)
) +
  plot_layout(widths = c(2, 2, 1), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre", axis.title = element_text(size = 18),
    legend.text = element_text(size = 18)
  )
suppressWarnings(print(p))
```

@@

If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (rAE for the surrogate compared to the ensemble `r overall_ae_median[model == "Surrogate", Mean] / overall_ae_median[model == "Ensemble", Mean] |> round(digits = 2)`). When using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (rAE `r overall_ae_median[model == "Surrogate", Median] / overall_ae_median[model == "Ensemble", Median] |> round(digits = 2)`). Across forecast horizons the same pattern of outperformance holds. However, the difference in relative performance was less than when the full probability distribution was accounted for with this becoming more marked as the forecast horizon increased (Figure \@ref(fig:assemble-eval) c).

@@

```{r assemble-eval, fig.height = 12, fig.width = 12, out.width = "95%", fig.cap = "Relative weighted interval score by location, horizon, and forecast date for the surrogate forecast model compared to the ensemble forecast model on the log scale. a.) The density of the relative score by horizon. Horizontal black lines give the 5\\%, 35\\%, 65\\%, and 95\\% quantiles. b.) The density of the relative score by month for a given forecast horizon stratified by the one and four-week forecast horizon. c.) The average relative weighted interval score and absolute error for the surrogate model compared to the ensemble forecast by forecast horizon. d.) The density of the relative score by forecast location stratified by the one and four-week forecast horizon. The dashed line on all plots indicates when the ensemble forecast is equivalent to the surrogate forecast. The vertical black lines on the y-axis give individual relative scores."}
p <- (
  (
    (
      plot_rwis_horizon / plot_rwis_by_month
    )  +
    plot_layout(heights = c(4, 5), guides = "collect")
  ) |
  (
    (
      plot_relative_summary(relative_summary) +
      plot_layout(guides = "keep")
    ) +
    plot_rwis_location +
    plot_layout(heights = c(1, 7), guides = "collect")
  )
) +
  plot_layout(guides = "collect", widths = c(1, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre", axis.title = element_text(size = 18),
    legend.text = element_text(size = 18)
  )
suppressWarnings(print(p))
```

@@

Stratifying calibration by quantile and forecast horizon the ensemble forecast was conservative at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
This behaviour became more prominent as the forecast horizon increased.
In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to over cover for intervals above the median.
At longer horizons, however, quantiles below the median became liable to under cover.

@@

```{r plot-coverage, fig.height = 12, fig.width = 18, out.width = "95%", fig.cap = "a.) Empirical coverage of the surrogate and ensemble forecasts at the 90\\%, 60\\%, and 30\\% prediction intervals stratified by forecast horizon. Ideally, a well-calibrated forecast should have empirical coverage for a given prediction interval that equals the nominal level of the interval (i.e., 30\\%, 60\\% and 90\\%, respectively). b.) Empirical coverage by quantile for both the surrogate and ensemble forecasts. A well-calibrated forecast should have empirical quantiles that match the theoretical ones. The green area of this figure corresponds to conservative forecasts. c.) Median relative weighted interval score by quantile and forecast horizon for the surrogate forecast compared to the ensemble forecast. d.) Bias of the ensemble and surrogate forecasts stratified by horizon."}
p <- (
  (
    (
      plot_coverage_range(coverage, ranges)
    )
  ) | (
    plot_coverage_quantiles(scores) + (
      plot_rel_score_by_quantile(relative_interval_score) +
      plot_bias(summary_performance)
    ) +
    plot_layout(heights = c(2, 1), guides = "collect")
  )
) +
  plot_layout(widths = c(1, 2), guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme(
    legend.position = "bottom", legend.margin = margin(),
    legend.justification = "centre", axis.title = element_text(size = 18),
    legend.text = element_text(size = 18)
  )
suppressWarnings(print(p))
```
